---
title: "Homework - features and elements of multiple regression"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

<hr>

# MVP


```{r}
library(tidyverse)
library(janitor)
library(fastDummies)
library(GGally)
library(ggfortify)
library(mosaic)
library(mosaicData)
library(leaflet)
library(raster)
```

1. Load the `housing_prices.csv` data set and undertake an initial exploration of the data. You will find details on the data set on the relevant [Kaggle page](https://www.kaggle.com/camnugent/california-housing-prices)

```{r}
housing_data <- read_csv(here::here("day2/data/housing_prices.csv"))
housing_data
```


2. We expect the `total_rooms` of houses to be strongly correlated with `total_bedrooms`. Use `ggpairs()` to investigate correlations between these two variables.

```{r}
housing_data %>% 
  select(total_rooms, total_bedrooms) %>% 
  ggpairs()


# yes strong positive correlation


```
 

3. So, we do find significant correlations. Let's drop `total_bedrooms` from the dataset, and use only `total_rooms` going forward.

```{r}
housing_data <- housing_data %>% 
  select(-total_bedrooms)
```



4. We are interested in developing a regression model for the `median_house_value` of a house in terms of the possible predictor variables in the dataset. 

  i. Use `ggpairs()` to investigate correlations between `median_house_value` and the predictors (this may take a while to run, don't worry, make coffee or something).

  ii. Perform further `ggplot` visualisations of any significant correlations you find.

```{r}
housing_data %>% 
  ggpairs()
```

```{r}
housing_data %>% 
  ggplot(aes(x = median_income, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

# strongest correlation with median house price is median income
```



5. Shortly we may try a regression model to fit the categorical predictor `ocean_proximity`. Investigate the level of `ocean_proximity` predictors. How many dummy variables do you expect to get from it?
    

```{r}
housing_data %>% 
  distinct(ocean_proximity) 

#4 with dummy encoding
```


6. Start with simple linear regression. Regress `median_house_value` on `median_income` and check the regression diagnostics.

```{r}
housing_data %>% 
  ggplot(aes(x = median_income, y = median_house_value)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

lm1 <- lm(median_house_value ~ median_income, housing_data)
summary(lm1)
plot(lm1)

# residuals vs fitted is fine 
# Normal - not good - std residuals are not normal
# Scale location - not good - heteroscedasticity 
```


7. Add another predictor of your choice. Check your assumptions, diagnostics, and interpret the model.



```{r}
lm2 <- lm(median_house_value ~ median_income + total_rooms, housing_data)
summary(lm2)
plot(lm2)

# adding total rooms made virtually no difference 
```


```{r}
housing_data_ocean <- housing_data %>% 
  select(median_house_value, median_income, ocean_proximity) %>% 
  dummy_cols(select_columns = "ocean_proximity",
                          remove_first_dummy = TRUE,
                          remove_selected_columns = TRUE) %>% 
  clean_names() %>% 
  mutate(across(starts_with("ocean"), as.logical)) 

lm2 <- lm(median_house_value ~ median_income + ocean_proximity_near_ocean + ocean_proximity_inland + ocean_proximity_island + ocean_proximity_near_bay, housing_data_ocean)
summary(lm2)
plot(lm2)

plotModel(lm2)

# same problems

```


# Extension
    
<br>

8. Try adding an interaction between `log(median_income)` and your chosen categorical predictor. Do you think this interaction term is statistically justified?

```{r}
lm2 <- lm(log(median_house_value) ~ log(median_income) + ocean_proximity_near_ocean + log(median_income):ocean_proximity_near_ocean, housing_data_ocean)
summary(lm2)
plot(lm2)

plotModel(lm2)

# its worse and still fails the diagnostics test

# is statistically significant for near ocean False, but nor for near Ocean TRUE
```


9. Find and plot an appropriate visualisation to show the effect of this interaction



```{r}
lm3 <- lm(log(median_house_value) ~ log(median_income) + ocean_proximity_near_ocean + ocean_proximity_inland + ocean_proximity_island + ocean_proximity_near_bay + log(median_income):ocean_proximity_near_ocean + log(median_income):ocean_proximity_inland + log(median_income):ocean_proximity_island + log(median_income):ocean_proximity_near_bay, housing_data_ocean)
summary(lm3)
plot(lm3)


```



```{r}
mod3 <- lm(log(median_house_value) ~ log(median_income) + ocean_proximity + log(median_income):ocean_proximity, data = housing_data)
summary(mod3)
```

```{r}
housing_data_2 <- housing_data %>% 
  mutate(rooms_per_household = total_rooms / households,
         bedrooms_per_room = total_bedrooms / total_rooms,
         people_per_house = population / households,
         bedrooms_per_person = total_bedrooms / population) %>% 
  select(median_house_value, median_income ,rooms_per_household, bedrooms_per_room, bedrooms_per_person, people_per_house, ocean_proximity)
  
  
ggpairs(housing_data_2)


```



```{r}
lm7 <- lm(median_house_value ~ median_income + bedrooms_per_room + ocean_proximity + median_income:bedrooms_per_room + median_income:ocean_proximity, housing_data_2)
summary(lm7)

plot(lm7)
```

```{r}
housing_data %>% 
  dplyr::select(longitude, latitude, housing_median_age) 


raster <- rasterFromXYZ(housing_data %>% 
  dplyr::select(longitude, latitude, median_house_value), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

plot(raster)

?rasterFromXYZ
```

```{r}


housing_data %>% 
  group_by(longitude, latitude) %>% 
  mutate(mean_median_house_value = mean(median_house_value))

#library(raster)
#library(leaflet)

raster <- rasterFromXYZ(housing_data %>% 
  dplyr::select(longitude, latitude, median_house_value), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

pal <- colorNumeric(c("blue", "red", "yellow"), values(raster),
  na.color = "transparent")

plot(raster)

leaflet() %>% 
  addTiles() %>% 
  addRasterImage(raster, colors = pal, opacity = 0.7) %>% 
  addLegend(pal = pal, values = values(raster),
    title = "Median House Value")



raster <- rasterFromXYZ(housing_data %>% 
  group_by(longitude, latitude) %>% 
  summarise(mean_median_house_value = mean(median_house_value)) %>%  
  dplyr::select(longitude, latitude, mean_median_house_value), crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

pal <- colorNumeric(c("blue", "red", "yellow"), values(raster),
  na.color = "transparent")

plot(raster)

leaflet() %>% 
  addTiles() %>% 
  addRasterImage(raster, colors = pal, opacity = 0.7) %>% 
  addLegend(pal = pal, values = values(raster),
    title = "Median House Value")


```

```{r}
lm7 <- lm(median_house_value ~ ocean_proximity + median_income, housing_data)
summary(lm7)
plotModel(lm7)

lm8 <- lm(median_house_value ~ median_income +  ocean_proximity, housing_data)
summary(lm8)
plotModel(lm8)
```

```{r}
df <- data.frame( x = rep( 0:1, each=4 ),
                  y = rep( 0:1,  4),
                  l = rnorm( 4 ))
df
dfr <- rasterFromXYZ(df)  #Convert first two columns as lon-lat and third as value                
 plot(dfr)
```

