---
title: "Introduction to `ggplot2` homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

```{r}
library(tidyverse)
library(CodeClanData)
```
```{r}
jamie_theme <- theme(panel.background = element_rect(fill = "white"),
                     panel.grid = element_line(colour = "#ededed"),
                     strip.background = element_rect(fill = "white")
                        ) 
```



# Homework Info

This homework recaps what you learnt in class today. 

You'll be looking at the `backpack` dataset contained within the `CodeClanData` package using tidyverse. You'll need to load the appropriate packages first.In all plots you make, ensure your labels are appropriate 

**You must write down a short summary of what each plot tells you in order to meet MVP.** Being an analyst is not just the ability to write code, you need to be able to interpret the visualisations and translate the relationships found (or not found) into plain language to explain to non-technical decision makers. Getting as much practice of this as you can while on this course is just (if not more!) important than the code!



# MVP

<br>

**Question 1.**  

Load in the libraries you need, and look at the `backpack` data. Take note of variable names and dimensions. 

```{r}
backpack %>% 
  names()

backpack %>% 
  dim()
```



**Question 2.**

What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution. 


```{r}
backpack %>% 
  ggplot(aes(backpack_weight)) +
  geom_histogram(binwidth = 2) +
  scale_x_continuous(breaks = seq(0, 40, 2)) +
  labs(title = "Backpack Weight Counts") +
  jamie_theme


backpack %>% 
  summarise(mean_backpack_weight = mean(backpack_weight))

# most of the students are distributed around the mean weight fo 11.66?kg/lbs?
```



**Question 3.**  

Are there differences in backpack weight between sex?  

```{r}
backpack %>% 
  ggplot(aes(backpack_weight, group = sex ,fill = sex)) +
  geom_histogram(binwidth = 2, position = "identity") +
  scale_x_continuous(breaks = seq(0, 40, 2)) +
  labs(title = "Backpack Weight Counts for Each Sex") +
  jamie_theme

backpack %>% 
  ggplot(aes(backpack_weight, group = sex ,fill = sex)) +
  geom_histogram(binwidth = 2) +
  scale_x_continuous(breaks = seq(0, 40, 2)) +
  facet_wrap(~ sex, nrow = 2) +
  labs(title = "Backpack Weight Counts for Each Sex") +
  jamie_theme


backpack %>% 
  group_by(sex) %>% 
  summarise(mean_backpack_weight = mean(backpack_weight))

# the distribution of male backpack weights is slightly heavier than female backpack weights
```



**Question 4.**  

What is the relationship between body weight and backpack weight?

```{r}
backpack %>% 
  ggplot(aes(x = body_weight, y = backpack_weight)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, show.legend = FALSE) +
  labs(title = "Body Weight vs Backpack Weight") +
  jamie_theme

# the geom smooth lines says that backback weight increases with body weight. Although I looking at the plot i would say that based on the data there is no relathionship 
  
```



**Question 5.**  

Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.  

```{r}
backpack %>% 
  ggplot(aes(x = body_weight, y = backpack_weight, colour = sex)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, show.legend = FALSE)

# the gradient of the "relationship" is slightly different. But again i'd say that there is no relathionship
```


**Question 6.**  

What **proportion** of students report back pain?   

```{r}
backpack %>% 
  count(back_problems) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>% 
  ggplot(aes(x = "students", y = n, fill = as.factor(back_problems))) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems") +
  guides(fill = guide_legend(title= "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

# proportion of kids with back problems is 0.32 (32%). 
```



**Question 7.**  

Is this proportion different between men and women? 

```{r}
backpack %>% 
  count(back_problems, sex) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>% 
  ggplot(aes(x = sex, y = n, fill = as.factor(back_problems))) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems") +
  guides(fill = guide_legend(title= "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))



# yes a higher proportion of females (approx ~ 0.44) have back pain compared to males (approx ~ 0.18)

```



**Question 8.**  

Now you've investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain.

Try both `backpack_weight` and `body_weight` as well as the `ratio`.   

*Hint: faceting would be useful here*.   

```{r}
backpack %>% 
  ggplot(aes(backpack_weight, group = back_problems, fill = as.factor(back_problems))) +
  geom_histogram(binwidth = 2) +
  scale_x_continuous(breaks = seq(0, 40, 2)) +
  labs(title = "Backpack Weight Counts") +
  jamie_theme

backpack %>% 
  ggplot(aes(body_weight, group = back_problems, fill = as.factor(back_problems))) +
  geom_histogram(binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 40, 2)) +
  labs(title = "Backpack Weight Counts") +
  jamie_theme

backpack %>% 
  ggplot(aes(body_weight, group = back_problems, fill = as.factor(back_problems))) +
  geom_histogram(binwidth = 10) +
  scale_x_continuous(breaks = seq(0, 40, 2)) +
  labs(title = "Backpack Weight Counts") +
  jamie_theme +
  facet_wrap(~ sex)

```



```{r}
backpack %>%
  group_by(sex) %>% 
  mutate(backpack_weight_cat = cut_number(
    x = backpack_weight,
    n = 4,
    labels = c("Quart1", "Quart2", "Quart3", "Quart4")
  )) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>%
  count(back_problems, backpack_weight_cat) %>%
  ggplot(aes(
    x = backpack_weight_cat,
    y = n,
    fill = as.factor(back_problems)
  )) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems\nIn each backpack weight quartile ") +
  guides(fill = guide_legend(title = "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  facet_wrap(~ sex)

backpack %>%
  mutate(backpack_weight_cat = cut_number(
    x = backpack_weight,
    n = 4,
    labels = c("Quart1", "Quart2", "Quart3", "Quart4")
  )) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>%
  count(back_problems, backpack_weight_cat) %>%
  ggplot(aes(
    x = backpack_weight_cat,
    y = n,
    fill = as.factor(back_problems)
  )) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems\nIn each backpack weight quartile ") +
  guides(fill = guide_legend(title = "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) 

# when sexes are combined clear trend of back pain increasing with increasing bag weight. not quite so clear when split be sex
  


  
```


```{r}
backpack %>%
  group_by(sex) %>% 
  mutate(body_weight_cat = cut_number(
    x = body_weight,
    n = 4,
    labels = c("Quart1", "Quart2", "Quart3", "Quart4")
  )) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>%
  count(back_problems, body_weight_cat) %>%
  ggplot(aes(
    x = body_weight_cat,
    y = n,
    fill = as.factor(back_problems)
  )) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems\nIn each body weight quartile ") +
  guides(fill = guide_legend(title = "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  facet_wrap(~ sex)

backpack %>%
  mutate(body_weight_cat = cut_number(
    x = body_weight,
    n = 4,
    labels = c("Quart1", "Quart2", "Quart3", "Quart4")
  )) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>%
  count(back_problems, body_weight_cat) %>%
  ggplot(aes(
    x = body_weight_cat,
    y = n,
    fill = as.factor(back_problems)
  )) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems\nIn each body weight quartile ") +
  guides(fill = guide_legend(title = "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) 
# when sexes are combined clear trend of back pain decreasing with increasing body weight. Different when broken down by sex
  
```


```{r}
backpack %>%
  mutate(
    bag_to_body_weight_ratio = backpack_weight / body_weight,
    ratio_cat = cut_number(
      x = ratio,
      n = 4,
      labels = c("Quart1", "Quart2", "Quart3", "Quart4")
    )
  ) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>%
  count(back_problems, ratio_cat) %>%
  ggplot(aes(
    x = ratio_cat,
    y = n,
    fill = as.factor(back_problems)
  )) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems\nIn each backpack weight to body weight ratio quartile ") +
  guides(fill = guide_legend(title = "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1))

backpack %>%
  group_by(sex) %>% 
  mutate(
    bag_to_body_weight_ratio = backpack_weight / body_weight,
    ratio_cat = cut_number(
      x = ratio,
      n = 4,
      labels = c("Quart1", "Quart2", "Quart3", "Quart4")
    )
  ) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>%
  count(back_problems, ratio_cat) %>%
  ggplot(aes(
    x = ratio_cat,
    y = n,
    fill = as.factor(back_problems)
  )) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems\nIn each backpack weight to body weight ratio quartile") +
  guides(fill = guide_legend(title = "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  facet_wrap(~ sex)

# when sexes are combined clear trend of back pai increaseing when the ratio of bag weight to body weight increase. Little bit fuzzier when broken down by sex but general trend similar
```

 

# Extensions

<br>

**Question 1.** 

Does reported backpain differ across the year groups? Ignore year 0 in your data. 

```{r}
backpack %>% 
  filter(year != 0) %>% 
  count(back_problems, year) %>%
  mutate(back_problems = case_when(back_problems == 1 ~ "Yes",
                                   back_problems == 0 ~ "No")) %>% 
  ggplot(aes(x = year, y = n, fill = as.factor(back_problems))) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems") +
  guides(fill = guide_legend(title= "Back Problems")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(1, 6, 1))

# yes the rates of back pain differ from year to year
```



**Question 2.** 

Is there a difference between how heavy the bags those in different years carry?


```{r}
backpack %>%
  filter(year != 0) %>% 
  mutate(backpack_weight_cat = cut_number(
    x = backpack_weight,
    n = 4,
    labels = c("Quart1", "Quart2", "Quart3", "Quart4")
  )) %>%
  count(year, backpack_weight_cat) %>%
  ggplot(aes(
    x = year,
    y = n,
    fill = as.factor(backpack_weight_cat)
  )) +
  geom_col(position = "fill") +
  jamie_theme +
  labs(x = "", y = "Proportion", title = "Proportion of Students with/without Back Problems\nIn each backpack weight quartile ") +
  guides(fill = guide_legend(title = "Bag weight quartile")) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  scale_x_continuous(breaks = seq(1, 6, 1))
# yes there is a difference - a high proportion of year sixes are carrying light bags
# year 5 carrying the highest proportion of v heavy bags
# year 2 are like cart horses have a high proprtion of high and v high but low proportion of back problems

```

