---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(broom)
library(janitor)
library(factoextra)
```

```{r}
customers <- read_csv(here::here("mall_customers.csv")) %>% clean_names()
```
You have been provided some customer data from a shopping centre. Do some exploratory analysis on the data. Comment on findings.

We are interested in creating a marketing campaign to target customers based on their spending score and annual income. Perform a k-means clustering to find if there are meaningful clusters in the data to target the customers.

Perform k-means clustering and chose a value of k.

Visualise the clustering for your chosen value of k.

Do you think the clustering seems a good fit for this data?

Comment on the attributes on one or two of the clusters (maybe even give them a label if you like - like in section 4.1 of the ‘Segmentation & clustering intro’ lesson).


```{r}
customers_subset <- customers %>% 
  select(annual_income_k, spending_score_1_100)

customers_subset %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100)) +
  geom_point()

# looks like there may be 5 clusters...
```

```{r}
customers_scaled <- customers_subset %>% 
  mutate(across(everything(), scale))
```

```{r}
fviz_nbclust(customers_scaled,
             kmeans,
             method = "wss",
             nstart = 25,
             k.max = 25)


fviz_nbclust(customers_scaled,
             kmeans,
             method = "silhouette",
             nstart = 25,
             k.max = 25,
             iter.max = 20)

fviz_nbclust(customers_scaled,
             kmeans,
             method = "gap",
             nstart = 25,
             iter.max = 20)

# lets go with five seems good on elbow and silhoutte
```

```{r}

clusters <- kmeans(customers_scaled,
       centers = 5,
       nstart = 25)

```

```{r}
customers_subset %>% 
  mutate(cluster = clusters$cluster) %>% 
  ggplot(aes(x = annual_income_k, y = spending_score_1_100, colour = as.factor(cluster))) +
  geom_point()
```


five groups
- low income - low spending score - on a budget
- low income - high spending score - reckless
- mid income - mid spending score - Goldilocks
- high income - low spending score - tight
- high income - high spending score - extravagant 
