---
title: "Simple linear regression homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center')
```

```{r}
library(ggfortify)
library(infer)
library(broom)
library(tidyverse)
library(modelr)

```

# MVP

The file `project_management.csv` contains data sampled from the recent work schedule of a small construction company. Column `estimated_length` contains the estimated length of a building job in days, while column `actual_length` contains the actual recorded length of the job in days. 

We are interested in determining the accuracy of the job estimations made by the company using simple linear regression, so we will eventually want to run a **simple linear regression** using `actual_length` as the dependent variable, and `estimated_length` as the independent variable.



* Load the data into a dataframe `project`

```{r}
project <- read_csv(here::here("data/project_management.csv"))
```



* Plot the data, taking `estimated_length` as the independent variable and `actual_length` as the dependent variable. 

```{r}
project %>% 
  ggplot() +
  geom_point(aes(x = estimated_length, y = actual_length))
```


* Calculate the correlation coefficient of `estimated_length` and `actual_length` and interpret the value you obtain.

```{r}
project %>% 
  summarise(cor(estimated_length, actual_length))

# indicates avery strong positive correlation between estimated length and actual length
```


* Perform a simple linear regression using `actual_length` as the dependent variable, and `estimated_length` as the independent variable. Save the model object to a variable. 

```{r}
lm1 <- lm(formula = actual_length ~ estimated_length, data = project)
lm1
```


* Interpret the regression coefficient of `estimated_length` (i.e. slope, gradient) you obtain from the model. How do you interpret the $r^2$ value reported by the model?

```{r}
glance(lm1)

# 64.7% of the variability of actual length can be predicted from variation in estimated length
```


* Is the relationship statistically significant? Remember, to assess this you need to check the $p$-value of the regression coefficient (or slope/gradient). But you should first check the regression diagnostic plots to see if the $p$-value will be reliable (don't worry about any outlier points you see in the diagnostic plots, we'll return to them in the extension).


```{r}
autoplot(lm1)

tidy(lm1)

# exluding outliers the diagnostic plots look fine.

# from p value (very close to 0) relationship can be defined as statistically significant
```



# Extension - Residuals vs Leverage

* Read [this material](https://boostedml.com/2019/03/linear-regression-plots-residuals-vs-leverage.html) on the **leverage** of points in regression, and how to interpret the `Residuals vs Leverage` diagnostic plot produced by plotting the `lm()` model object. So far we've been using the `autoplot()` function to plot the model objects produced by `lm()`, but you can see the base `R` equivalent by doing something like `plot(model)` where `model` is an `lm()` object.

```{r}
plot(lm1)
```


* Return to your plot from earlier, and now label the data points with their row number in the data frame using `geom_text()` [**Hint** - you can pass `aes(label = 1:nrow(project))` to this layer to generate row index labels]
  - Identify by eye any points you think might be outliers and note their labels.
  - Further split your outliers into those you think are 'influential' or 'non-influential' based on a visual assessment of their leverage.


```{r}
project %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_text(aes(label = 1:nrow(project))) +
  geom_smooth(method = "lm")

# biggest outliers by eye are 5 and 18. 5 influencial, 18 not influencial
```





* Use your model object from earlier and confirm your visual assessment of which points are 'influential' or 'non-influential' outliers based on Cook's distance. You can get a useful plot of Cook's distance by passing argument `which = 4` to `autoplot()`. Or try the base `R` `plot()` function for comparison [e.g. `plot(model)`; you can also use `par(mfrow = c(2,2))` just before the `plot()` command to get a nice two-by-two display]!
```{r}
plot(lm1)

# confirmed that 5 is influential

# 31 also an outlier
```

* Obtain the intercept and regression coefficient of variable `estimated_length` for a simple linear model fitted to data **omitting one of your non-influential outlier points**. 
  - How different are the intercept and coefficient from those obtained above by fitting the full data set? Does this support classifying the omitted point as non-influential? 
  - Plot the data points, this regression line and the regression line for the full data set. How different are the lines?
  
```{r}
lm1 <- lm(formula = actual_length ~ estimated_length, data = project)
lm1
a <- coefficients(lm1)



lm_non_inf <- project %>% filter(!row_number() == 18) %>% lm(formula = actual_length ~ estimated_length, .)
lm_non_inf
b <- coefficients(lm_non_inf)

project %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_abline(slope = a[2], a[1]) +
  geom_abline(slope = b[2], b[1], colour = "red")

# hardly any difference to the gradient, minor difference to the intercept
```


* Repeat the procedure above, but this time **omitting one of your influential outliers**. 

```{r}
lm1 <- lm(formula = actual_length ~ estimated_length, data = project)
lm1
a <- coefficients(lm1)



lm_inf <- project %>% filter(!row_number() == 5) %>% lm(formula = actual_length ~ estimated_length, .)
lm_inf
c <- coefficients(lm_inf)

project %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_abline(slope = a[2],intercept =  a[1]) +
  geom_abline(slope = c[2], intercept =  c[1], colour = "red")


project %>% 
  add_predictions(lm1, var = "pred1") %>% 
  add_predictions(lm_inf, var = "pred2") %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point() +
  geom_line(aes(y = pred1), col = "black") +
  geom_line(aes(y = pred2), col = "red") 

# much bigger difference in both values. Line is very different also
```

# Additional resources

* There are various techniques to perform what is known as 'robust regression' on a dataset. Robust methods are less affected by the presence of outliers. See the `rlm()` function ('robust linear model') in the `MASS` package and [this](http://www.alastairsanderson.com/R/tutorials/robust-regression-in-R/) blog post.



