---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(e1071)
library(skimr)
library(janitor)
library(infer)
library(fastGraph)
```

Task 1.
Load the data again, clean_names(), and re-familiarise yourself with it

```{r}
ames_data <- read_csv(here::here("data/ames.csv")) %>% 
  clean_names()
ames_data
```


Task 2.
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?


```{r}
ames_data %>% 
  ggplot() +
  geom_histogram(aes(x = lot_area, y = ..density..), binwidth = 1000) +
  stat_function(
    fun = dnorm,
    args = list(
      mean = mean(ames_data$lot_area),
      sd = sd(ames_data$lot_area)
    )
  )

# distribution is not normal there is a right skew with outliers with large values of lot_area. Peak values also much higher than expected
```



Task 3.
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.

```{r}

infer_resample <- ames_data %>% 
  specify(response = lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

infer_resample %>% 
  visualise(bins = 30) 

# looks normal

## weird code 

infer_resample <- ames_data %>% 
  specify(response = lot_area) %>% 
  hypothesize(null = "point", mu = 40) %>%
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "t")

obs_stat <- ames_data %>%
  specify(response = lot_area) %>%
  hypothesize(null = "point", mu = 40) %>%
  calculate(stat = "t")

visualize(infer_resample, method = "both", bins = 30)

  



```


Task 4.
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution


```{r}
infer_ci95 <- infer_resample %>% 
  get_ci(level = 0.95, type = "percentile")
  

infer_resample %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci95)
```


Task 5.
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?

```{r}
infer_ci99 <- infer_resample %>% 
  get_ci(level = 0.99, type = "percentile")
  

infer_resample %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci99)

# it is broader - yes this makes sense
```



Task 6.
Calculate the point estimate of the mean(lot_area)

```{r}
# boot strap mean

infer_resample %>% 
  summarise(mean = mean(stat))

# original sample mean

ames_data %>% 
  summarise(mean = mean(lot_area))

# the means are very similar which is expected 
```



Task 1.
Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? [Investigate reps from 200 up to 50000, memory of your laptop permitting].

```{r}
#5000

infer_before_1920 <- ames_data %>% 
  select(year_built) %>% 
  mutate(before_1920 = if_else(year_built < 1920, "Yes", "No")) %>% 
  specify(response = before_1920, success = "Yes") %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "prop")

infer_ci95 <- infer_before_1920 %>% 
  get_ci(level = 0.95, type = "percentile")
infer_ci95

infer_before_1920 %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci95)

#point estimate

infer_before_1920 %>% 
  summarise(mean = mean(stat))


```

```{r}
#200

infer_before_1920 <- ames_data %>% 
  select(year_built) %>% 
  mutate(before_1920 = if_else(year_built < 1920, "Yes", "No")) %>% 
  specify(response = before_1920, success = "Yes") %>% 
  generate(reps = 200, type = "bootstrap") %>% 
  calculate(stat = "prop")

infer_ci95 <- infer_before_1920 %>% 
  get_ci(level = 0.95, type = "percentile")
infer_ci95

infer_before_1920 %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci95)

#point estimate

infer_before_1920 %>% 
  summarise(mean = mean(stat))
```


```{r}
#20000

infer_before_1920 <- ames_data %>% 
  select(year_built) %>% 
  mutate(before_1920 = if_else(year_built < 1920, "Yes", "No")) %>% 
  specify(response = before_1920, success = "Yes") %>% 
  generate(reps = 20000, type = "bootstrap") %>% 
  calculate(stat = "prop")

infer_ci95 <- infer_before_1920 %>% 
  get_ci(level = 0.95, type = "percentile")
infer_ci95

infer_before_1920 %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci95)

#point estimate

infer_before_1920 %>% 
  summarise(mean = mean(stat))


# distributions look nicer with more reps but cis dont seem much different 
```

